---
title: "Minizinc Interface for R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Minizinc Interface in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
Minizinc is a high-level constraint modeling language that researchers and developers can use to describe their domain problems in. The biggest advantage of using Minizinc is that it is solver-independent and provides a unified interface to many popular constraint satisfaction and optimization solvers.

The goal of this package is to bring the power of Minizinc to R users in a way that does not require any knowledge of Minizinc. 
This way, users can program in R like they would normally do, and still have the access to all features and constraints provided by Minizinc as well as low-level solvers. 

This document introduces you to the basic usage of the package. 

Most of the package is written in object-oriented style using R6. If you need a refresher on R6 OOP system or are not familiar with it, here is a basic [tutorial](https://adv-r.hadley.nz/r6.html).

For more information on Minizinc, visit [Minizinc](https://www.minizinc.org/). 

A lot of information in this vignette is taken from the official [Minizinc Handbook](https://www.minizinc.org/doc-2.4.3/en/modelling.html). 

# Variables
There are two types of variables in interface package -- decision and parameter. 

Parameter variable is a variable that is used to specify a parameter in a given problem. For example, a maximum value a decision variable can hold. Parameter variables can be integers (int), floating point numbers (float), Booleans (bool) and strings (string). 

Decision variables are the unknowns that Minizinc model is finding solutions for. We do not need to give them a value, but instead we give them a domain of possible values. Decision variables need to satisfy a set of constraints that a user specifies in the model. 
Below are examples of how we can create both variable types. 

```{r}
library(minizinc)
par = Variable$new(type = "float", kind = "parameter", value = 0.8)
dec = Variable$new(type = "float", kind = "decision", domain = c(0, 1))
```

More information on variable is available in `?Variable`. 

# Constraints
Constraints specify the Boolean expressions that decision variables must satisfy. Examples of such constraints are given below. 

```{r}
par = Variable$new(type = "float", kind = "parameter", value = 0.8)
var1 = Variable$new(type = "float", kind = "decision", domain = c(0, par$value))
var2 = Variable$new(type = "float", kind = "decision", domain = c(0, par$value))

constraint = Constraint$new(constraint = ">=", variables = c(var1, var2))
```

For now, this package can only support relational constraints that involve two variables at a time only. 

# Solvers
A standard Minizinc distribution contains a few solvers. To find a solution to problem, we can choose which solver for a given problem. The default solver is Gecode. In order to use a solver, we need to first create a solver object in R.

```{r}
s = Solver$new(name = "gecode")
```

There is also `config` field in Solver class that is not currently utilized. Support for manually configuring solvers will be added later. 

# Models
Now that we have all parts to specify a problem, we need to put these parts together in a model object. We also need to specify an objective of our problem (satisfy, maximize, minimize). 

Below is an example of map coloring problem of Australian states. 

```{r}
# parameter variable
p1 = Variable$new(type = "int", value = 3, kind = "parameter")

# decision variables
v1 = Variable$new(type = "int", kind = "decision", domain = c(1, p1$value))
v2 = Variable$new(type = "int", kind = "decision", domain = c(1, p1$value))
v3 = Variable$new(type = "int", kind = "decision", domain = c(1, p1$value))
v4 = Variable$new(type = "int", kind = "decision", domain = c(1, p1$value))
v5 = Variable$new(type = "int", kind = "decision", domain = c(1, p1$value))
v6 = Variable$new(type = "int", kind = "decision", domain = c(1, p1$value))
v7 = Variable$new(type = "int", kind = "decision", domain = c(1, p1$value))
vars = c(v1, v2, v3, v4, v5, v6, v7)

# constraints
c1 = Constraint$new(constraint = "!=", variables = c(v1, v2))
c2 = Constraint$new(constraint = "!=", variables = c(v1, v3))
c3 = Constraint$new(constraint = "!=", variables = c(v2, v3))
c4 = Constraint$new(constraint = "!=", variables = c(v2, v4))
c5 = Constraint$new(constraint = "!=", variables = c(v3, v4))
c6 = Constraint$new(constraint = "!=", variables = c(v3, v5))
c7 = Constraint$new(constraint = "!=", variables = c(v3, v6))
c8 = Constraint$new(constraint = "!=", variables = c(v4, v5))
c9 = Constraint$new(constraint = "!=", variables = c(v5, v6))
constr = c(c1, c2, c3, c4, c5, c6, c7, c8, c9)

# solver
s = Solver$new(name  = "gecode")

# putting it all together
m = Model$new(parameter = c(p1), decision = vars, constraints = constr, 
              objective = "satisfy")
```

Before solving the problem we also need to specify the path to our Minizinc installation. For example on my system, 

```{r}
set_path("/home/damir/software/MiniZincIDE-2.4.2-bundle-linux/bin/minizinc")
```


After everything is set up, all we need to do is 

```{r}
res = solve(m, s)
print(res)
```

A Minizinc implementation of the same problem is given below:

```{minizinc}
% Colouring Australia using nc colours
int: nc = 3;

var 1..nc: wa;   var 1..nc: nt;  var 1..nc: sa;   var 1..nc: q;
var 1..nc: nsw;  var 1..nc: v;   var 1..nc: t;

constraint wa != nt;
constraint wa != sa;
constraint nt != sa;
constraint nt != q;
constraint sa != q;
constraint sa != nsw;
constraint sa != v;
constraint q != nsw;
constraint nsw != v;
solve satisfy;

output ["wa=\(wa)\t nt=\(nt)\t sa=\(sa)\n",
        "q=\(q)\t nsw=\(nsw)\t v=\(v)\n",
         "t=", show(t),  "\n"];
```


That is. We just solved our first constraint satisfaction problem without knowing anything about Minizinc and just programming in R!
